cardinalR: Generating interesting high-dimensional data structures

### 2025-11-20 paper-cardinalR.Rmd
Rewrote longlinear paragraph

The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise. Each variable is formed as $X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i$, where $\text{scale}\_i \sim U(-10, 10)$ determines the orientation of the line in each dimension, $\text{shift}\_i \sim U(-300, 300)$ offsets the line to separate dimensions, and $\epsilon \sim N(0, (0.03n)^2)$ introduces Gaussian noise.

>>>>

<!-- The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise. Each variable is formed as $X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i$, where $\text{scale}\_i \sim U(-10, 10)$ determines the orientation of the line in each dimension, $\text{shift}\_i \sim U(-300, 300)$ offsets the line to separate dimensions, and $\epsilon \sim N(0, (0.03n)^2)$ introduces Gaussian noise.  -->

The `gen_longlinear(n, p)` function generates a high-dimensional dataset representing a long linear structure with noise. Each variable is constructed by first creating a sequence of integers from $0$ to $n-1$, adding Gaussian noise with standard deviation $0.03n$, and then applying a random scaling and shift. Specifically, for the $j$-th dimension, a scale factor $\text{scale}_j \sim U(-10, 10)$ determines the orientation and stretching of the line, while a shift factor $\text{shift}_j \sim U(-300, 300)$ is added inside the multiplication to offset the line. This results in each variable being of the form $X_j = \text{scale}_j \cdot \big((0,1,\dots,n-1) + \epsilon + \text{shift}_j\big)$, where $\epsilon \sim N(0, (0.03n)^2)$ is independent Gaussian noise.

### 2025-11-20 paper-cardinalR.tex
Rewrote longlinear paragraph

The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes (\url{https://jayanilakshika.github.io/cardinalR/reference/index.html}), which can then be combined into a single dataset including scaling, rotation, and translation. The package is available on CRAN, and the source is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.

@@

The \texttt{gen\_longlinear(n,\ p)} function generates a high-dimensional dataset representing a long linear structure with noise. Each variable is formed as \(X_i = \text{scale}_i \cdot (0,1,\dots,n{-}1 + \epsilon) + \text{shift}_i\), where \(\text{scale}\_i \sim U(-10, 10)\) determines the orientation of the line in each dimension, \(\text{shift}\_i \sim U(-300, 300)\) offsets the line to separate dimensions, and \(\epsilon \sim N(0, (0.03n)^2)\) introduces Gaussian noise.

>>>>

The \texttt{cardinalR} package is built on a modular framework where individual geometric generators (e.g., Gaussian, cone, sphere) create well-defined shapes (A full list of available shape generators are available at \url{https://jayanilakshika.github.io/cardinalR/reference/index.html}.), which can then be combined into a single dataset including scaling, rotation, and translation. The package is available on CRAN, and the source is available on GitHub at \href{https://github.com/JayaniLakshika/cardinalR}{JayaniLakshika/cardinalR}.

@@

The \texttt{gen\_longlinear(n,\ p)} function generates a high-dimensional dataset representing a long linear structure with noise. Each variable is constructed by first creating a sequence of integers from \(0\) to \(n-1\), adding Gaussian noise with standard deviation \(0.03n\), and then applying a random scaling and shift. Specifically, for the \(j\)-th dimension, a scale factor \(\text{scale}_j \sim U(-10, 10)\) determines the orientation and stretching of the line, while a shift factor \(\text{shift}_j \sim U(-300, 300)\) is added inside the multiplication to offset the line. This results in each variable being of the form \(X_j = \text{scale}_j \cdot \big((0,1,\dots,n-1) + \epsilon + \text{shift}_j\big)\), where \(\epsilon \sim N(0, (0.03n)^2)\) is independent Gaussian noise.
